\section{Methodology}\label{s:methodology}

\subsection{Existing metrics of planetary wave activity}\label{s:existing_metrics}

In analysing planetary wave activity in the Southern Hemisphere (i.e. the ZW1 and/or ZW3 patterns), previous studies have tended to define metrics based on either grid point values or Fourier decomposition. With respect to the former, \citet{Raphael2004} define a ZW3 index that is essentially the average 500 hPa geopotential height zonal anomaly across three key points (the annual average location of the ridges of the ZW3 pattern in the 500 hPa geopotential height field). This metric was applied in a subsequent study of Antarctic sea ice variability \citep{Raphael2007} and is attractive due to its simplicity, however its stationary nature means that it cannot fully capture the subtle \citep[but not insignificant at around 15 degrees of longitude on average;][]{vanLoon1984,Mo1985} seasonal migration in the phase of the ZW3 or the occurence of patterns whose phase is not close to average.

\citet{Hobbs2007} and \citet{Hobbs2010} use a Fourier Transform to express the upper tropospheric geopotential height in the frequency domain as opposed to the spatial domain. The output of a Fourier Transform is essentially a magnitude and phase coefficient for each wavenumber (or frequency/harmonic - the terminology differs in the literature), so those studies simply picked out the coefficients corresponding to the ZW1 (and ZW3 in the case of \citep{Hobbs2010}). While this is an improvement on the grid point method in the sense that the phase is allowed to vary, a shortcoming is that the result is a constant amplitude wave over the entire domain. The two major anticyclones associated with the ZW3 pattern (located near New Zealand and South America respectively) are known \citet{Hobbs2010} to be positively covariant with respect to their location (indicating a coordinated wave pattern) but not amplitude, while in many cases ZW1- and/or ZW3-like variability is only prevalent over part of the hemisphere. As discussed in the seminal work of \citet{vanLoon1972}, it is clear that the other Fourier components (i.e. the non-wave 1 or 3 coefficients) are required to modulate the amplitude of the ZW1 and ZW3 variability, and potentially vital information is lost if those extra components are not incoproated when defining a metric of planetary wave activity. 

None of the aforementioned studies attempted to combine their ZW1 and ZW3 metrics to get a measure of total planetary wave activity, so for an example of this we must turn to the Northern Hemisphere. In analysing the relationship between planetary wave activity and regional weather extremes, \citet{Screen2014} calculated the magnitude coefficient (from a Fourier Transform of the 500 hPa geopotential height) for a range of wavenumbers of interest, and then simply counted the number of positive and negative anomalies. While this may be an appropriate appraoch for the Northern Hemisphere, it too fails to account for that fact that some of the waveforms in a Fourier Transform simply exist to modulate others (which is something that is noted by \citet{Screen2014}) and thus it may not be apporpriate to count those.  