\subsection{Analysis techniques}

\subsubsection{Anomalies}
All anomaly data discussed in the paper represent the daily anomaly. For instance, in preparing the 30-day running mean surface air temperature anomaly data series, a 30-day running mean was applied to the daily surface air temperature data and then the anomaly at each timestep was calculated by subtracting the mean value for that specific day over the entire study period.

\subsubsection{Composites}
Annual and seasonal mean composites are presented throughout the paper for key variables. To produce these composites, a mean spatial map was calculated across all timesteps that (a) fell within the season of interest, and (b) exceeded a specified threshold value. To determine the statistical significance of the composite mean at each grid point, two-sided, independent sample t-tests were used to calculate the probability ($p$) that the composite mean value was equal to the climatological (i.e. all timesteps) mean for that season.

\subsubsection{Normalization}
In order to aid visual comparison, a number of plots throughout the paper present the `normalized' values of an index. The normalization process involves subtracting the mean of the data series and then dividing by the standard deviation. This is also known as the standardized anomaly \citep{Wilks2011}. 
%Wilks (p 46) call this a standardised anomaly

\subsubsection{Periodograms} %From Wilks2011
The characteristics of data series that have been Fourier-transformed are often summarized using a plot known as a periodogram or Fourier line spectrum \citep{Wilks2011}. These plots are also referred to as a power or density spectrum, and most commonly display the squared amplitudes ($C_k^2$) of the Fourier Transform coefficients as a function of their corresponding frequencies ($\omega_k$). As an alternative to the squared amplitude, we have chosen to rescale the vertical axis and instead use the $R^2$ statistic commonly computed in regression analysis. The $R^2$ for the $k$th harmonic is simply:

\begin{equation}\label{eq:variance_explained}
R_k^2 = \frac{(n/2)C_k^2}{(n-1)s_y^2}
\end{equation}

where $s_y^2$ is the sample variance of the data series and $n$ the length of the data series. This rescaling is particularly useful as it shows the proportion of variance in the original data series accounted for by each harmonic \citep{Wilks2011}.

\subsubsection{Climate indices}
Two of the major modes of SH climate variability are the El Ni\~{n}o Southern Oscillation (ENSO) and Southern Annular Mode (SAM). In order to assess their relationship with the PWI, the Ni\~{n}o 3.4 index of \citet{Trenberth2001} and Antarctic Oscillation Index of \citet{Gong1999} were calculated from 30-day running mean data (i.e. the same timescale that was used to calculate the PWI). The former represents the sea surface temperature anomaly for a region in the central tropical Pacific Ocean (5$^{\circ}$S-5$^{\circ}$N and 190-240$^{\circ}$E), while the latter represents the difference of zonal mean sea level pressure between 40$^{\circ}$S and 65$^{\circ}$S. 

